---
title: "plots"
author: "Jesper Fischer Ehmsen"
date: "`r Sys.Date()`"
output:
  word_document: default
  html_document: default
---

This markdown document contains the code to generate the figures featured in the Manuscript.

<!-- Please note that the packages for making the excel tables from flextable "YesSiR" needs to be installed with the following commands: -->
<!-- install.packages("devtools") -->
<!-- devtools::install_github("Sebastien-Le/YesSiR") -->

```{r}
# Load necessary libraries

if (!require("pacman")) install.packages("pacman")
pacman::p_load("renv", "here", "knitr")

knitr::opts_chunk$set(echo = TRUE)

#renv::restore(project=here::here())

required_packages = c("tidyverse", "gamlss", "glmmTMB", "gghalves", "ggtext", "osfr", "scales","cowplot","flextable","ggeffects","grid","magick","patchwork","png",
"YesSiR","ggh4x","readr")

lapply(required_packages, library, character.only = TRUE)
detach("package:renv", unload = T)

```

```{r}
# Source additional scripts for plotting and utilities
source(here::here("scripts","plots.R"))
source(here::here("scripts","utils.R"))


# Get data
data <- get_data(rerun = FALSE)
df <- data$data

# Get HGF trajectories
trajfeel2 = get_hgf(df, rerun = F)
```

# Figure 1: Overview of experimental design and behavioural measures related to thermosensory learning

```{r, fig.height=7.2, fig.width=7.2}
plot1 = plot1(df)
plot1
ggsave(here::here("Figures","figure1.png"), plot = plot1, dpi = 600)
```

# Figure 2: Thermosensory ratings and impact of expectations on temperature and TGI perception

```{r, fig.height=7.2, fig.width=7.2}
plot2 = plot2(df)
plot2
ggsave(here::here("Figures","figure2.png"), plot = plot2, dpi = 600)
```

# Figure 3: Perceptual and response model

```{r, fig.height=7.2, fig.width=7.2}
source(here::here("scripts","plots.R"))
source(here::here("scripts","utils.R"))
plot3 = plot3(trajfeel2)
plot3
ggsave(here::here("Figures","figure3.png"), plot = plot3, dpi = 600)
```

# Figure 4

```{r, fig.height=7.2, fig.width=7.2}
source(here::here("scripts","plots.R"))

plot_4 = plot4_v2(trajfeel2)
plot_4
ggsave(here::here("Figures","figure4.png"), plot = plot_4, dpi = 600)
```

# Table 1: Computational correlates of MT weighted images:

```{r}
source(here::here("scripts","plots.R"))
source(here::here("scripts","utils.R"))

table1 = draw_table_main()
table1[[1]]
table1[[2]]

draw_table_supplementary()
```

# Figure 5: Computational neuroanatomy of the Thermal Grill Illusion

```{r, fig.height=7.2, fig.width=7.2}
# source(here::here("scripts","plots.R"))
# source(here::here("scripts","utils.R"))
# 
# plot5 = plot5(R = TRUE)
# plot5 
# ggsave(here::here("Figures","figure5.png"), plot = plot5, dpi = 600)
```

```{r,fig.height=6.4, fig.width=7.2}
# source(here::here("scripts","plots.R"))
# source(here::here("scripts","utils.R"))
# 
# plot5 = plot5_v2()
# plot5 
# ggsave(here::here("Figures","figure5.png"), plot = plot5, dpi = 600)

```


